---
title: "Housing Market - Seattle Area"
author: "Anti-Code Group"
date: "10/5/2019"
output:  
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(scipen = 999)
```

```{r basicfunct, include=FALSE}
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
```

```{r base_lib, include=FALSE}
loadPkg("dplyr")
loadPkg("tidyr")
loadPkg("tidyverse")
loadPkg("ggplot2")
loadPkg('ggmap')
loadPkg('rjson')
loadPkg('jsonlite')
loadPkg('leaflet')
loadPkg('grDevices')
loadPkg('scales')
loadPkg('RCurl')
loadPkg('sp')
loadPkg('geojsonio')
loadPkg('lmtest')
loadPkg("faraway")
loadPkg("corrplot")
loadPkg("modelr")
```

# Description of the Data

According to Kaggle: "this dataset contains house sale prices for King County, WA which includes Seattle. It includes homes sold between May 2014 and May 2015. The dataset has 21,613 observations and 21 variables. The data has a mean housing price of $540,088 and a standard deviation of $367,127.

## Source Data

https://www.kaggle.com/harlfoxem/housesalesprediction/data

https://github.com/matteobucalossi50/DATS6101_Project

```{r, echo=FALSE}
kc_house_data <- read.csv("kc_house_data.csv")
str(kc_house_data)
colnames(kc_house_data)
```

```{r clean, include=FALSE}
kc_house_data <- subset(kc_house_data, select = -c(9, 10))
kc_house_data <- subset(kc_house_data, kc_house_data$bedrooms != 0)
kc_house_data <- subset(kc_house_data, kc_house_data$bathrooms != 0)
kc_house_data <- subset(kc_house_data, kc_house_data$bedrooms < 30)
kc_house_data <-  drop_na(kc_house_data)
kc_house_data$condition <- as.factor(kc_house_data$condition)
kc_house_data$grade <- as.factor(kc_house_data$grade)
```
Explain more of grade, condition. Question with answer per section.
Use year renovated for tests and comparison.

## Summary for Price
```{r, echo=FALSE}
summary(kc_house_data$price)
print(mean(kc_house_data$price))
print(sd(kc_house_data$price))
print(var(kc_house_data$price))
```

## Geographic Coverage of Data
Below is a visualization of the points in the dataset by price, plotted with the leaflet library.
 Note that the data have been divided by quantiles to provide a better understanding of the distribution of rental price.
 
```{r geo, echo=FALSE}
price.bins <- c(0, 250000, 500000, 750000, 1000000, 1250000, 1500000, 1750000, 2000000, 8000000)
qpal <-  colorBin(palette = 'GnBu', kc_house_data$price, bins= price.bins, n = 9)
house.map <- leaflet(kc_house_data) %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addCircleMarkers(lng = ~long, lat = ~lat, 
                   stroke = FALSE, 
                   fillOpacity = 5, 
                   color = ~qpal(price), 
                   radius = 2,
                   label = ~as.character(paste0('Price: $', price, ", ", 'condition: ', condition, ", ", 'year built: ', yr_built))) %>%
  addLegend('bottomright', pal = qpal, values = ~price, opacity = 1, title = 'Price', labFormat = labelFormat(prefix = '$', between = '- $'))
house.map
```

# Exploratory Data Analysis

## Distribution of Price

The following are housing price distributions (histograms). The first includes the frequency of all price occurences, and the other only shows those less than or equal to $2M (the vast majority of data points are below this price point). Also included is a Q-Q plot showing the distribution of the data.

```{r price hist, echo=FALSE}
hist(kc_house_data$price, xaxt="n", ylim = c(0,14000), col = "green", main = "Housing Price Histogram", xlab = "Housing Price ($)", cex.axis = .75)
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))
hist(kc_house_data$price[kc_house_data$price<=2000000], xaxt="n", ylim = c(0,5000), col = heat.colors(20), main = "Housing Price Histogram, $0-$2M Only", xlab = "Housing Price ($)", cex.axis = .75)
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))
qqnorm(kc_house_data$price, pch = 20, cex = .5, main = "Housing Prices: Normal Q-Q Plot", ylab = "Housing Price ($)", cex.axis = .75)
```

##The following is a summary of price per square foot, per annum.

```{r price per year}
date_price <- kc_house_data[c('date', 'price', 'sqft_living', 'sqft_lot')]
date_price$date <- as.character(date_price$date)
date_price$date <- substr(date_price$date, 1, nchar(date_price$date)-11)
date_price <- date_price[order(date_price$date),]
date_price$date <- as.factor(date_price$date)
date_price %>%
  group_by(date) %>%
  summarise(price = sum(price), sqft_living=sum(sqft_living), sqft_lot=sum(sqft_lot), price_per_sqft = price/sqft_living)
```

price per sqft:  
  2014 -> 36.06  
  2015 -> 35.11  

  
## High observations

The following are a summary of the first few rows for bedrooms, land area, and living space (price showed is average for that variable level (e.g. $262,000 is the average price for grade 3 construction).

Top 5 rows of prices across our variable set.

```{r top price, echo=FALSE}
print(top_n(kc_house_data, 5, price)) %>% arrange(desc(price))
```

Top 5 rows of sqft. living across our variable set.

```{r top sqftliv, echo=FALSE}
print(top_n(kc_house_data, 5, sqft_living)) %>% arrange(desc(sqft_living))
```

Top 5 rows of land area across our variable set.

```{r top sqftlot, echo=FALSE}
print(top_n(kc_house_data, 5, sqft_lot))  %>% arrange(desc(sqft_lot))
```

Top 4 rows of bedroom #'s across our variable set.

```{r top bedrooms, echo=FALSE}
print(top_n(kc_house_data, 4, bedrooms))  %>% arrange(desc(bedrooms))
```

Mean prices for each condition level.

```{r top avg cond, echo=F}
print(aggregate(price~condition,data=kc_house_data,FUN=mean))
```

Mean prices for each grade level.

```{r rop avg grade, echo=F}
print(aggregate(price~grade,data=kc_house_data,FUN=mean))
```

Below is a scatterplot of price vs. living space, along with their correlation.

```{r plot sqft living, echo=FALSE}
plot(kc_house_data$sqft_living, kc_house_data$price, xaxt="n", yaxt="n", pch = 20, cex = .1, xlab = "Square Feet of Living Space", ylab = "Housing Price ($)", cex.axis = .75, main = "Housing Price vs. Living Space")
abline(lm(kc_house_data$price ~ kc_house_data$sqft_living, data = kc_house_data), col = "orange")
legend(x='topright', legend=paste('Correlation =',round(cor(kc_house_data$sqft_living, kc_house_data$price),5)))
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))
axis(side=2, at=axTicks(2), 
     labels=formatC(axTicks(2), format="d", big.mark=','))
```

## Boxplot & Scatterplot for Living Space

Below are a boxplot and scatterplot for living space vs. price.

```{r boxplot sqft living, echo=FALSE}
max(kc_house_data$sqft_living)
min(kc_house_data$sqft_living)
sqft.living <- cut(kc_house_data$sqft_living, breaks = c(289,2940,5590,8240,10890,13540), labels = c("290-2,940", "2,941-5,590", "5,591-8,240", "8,241-10,890", "10,891-13,540"))
ggplot(kc_house_data, aes(x=sqft.living, y=price, fill=sqft.living)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Living Space") + ylab("Housing Price ($)") + xlab("Square Feet of Living Space") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 
```


## Comparison price and lot sqft

The following includes histograms of lot square feet (including one equal to or below 25,000 sqft., where the vast majority of data points lie). Also included is a scatterplot showing the correlation between price and land area.

```{r lot summary, echo=FALSE}
summary(kc_house_data$sqft_lot)
```

```{r plot lot, echo=FALSE}
hist(kc_house_data$sqft_lot, col = "yellow", cex.axis = .75, xlab = "Land Area (sqft.)", main = "Histogram of Land Area")
hist(kc_house_data$sqft_lot[kc_house_data$sqft_lot<=25000], xaxt="n", cex.axis = .75, xlab = "Land Area (sqft.)", main = "Histogram of Land Area, up to 25,000 sqft.", col = heat.colors(13))
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))
plot(kc_house_data$sqft_lot,kc_house_data$price, cex.axis = .75, pch = 20, cex = .1, xlab = "Land Area (sqft.)", ylab = "Housing Price ($)", xaxt="n", yaxt="n", main = "Housing Price vs. Land Area")
abline(lm(kc_house_data$price~kc_house_data$sqft_lot, data = kc_house_data), col = "orange")
legend(x='topright', legend=paste('Correlation =',round(cor(kc_house_data$sqft_lot, kc_house_data$price),5)))
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))
axis(side=2, at=axTicks(2), 
     labels=formatC(axTicks(2), format="d", big.mark=','))
```

## Comparison of Land Area and Housing Price

Included below is a boxplot of land area vs. housing price.

```{r boxplot lot, echo=FALSE}
max(kc_house_data$sqft_lot)
min(kc_house_data$sqft_lot)
sqft.lot <- cut(kc_house_data$sqft_lot, breaks = c(519,330688,660856,991023,1321191, 1651359), labels = c("520-330K", "330K-660K", "660K-991K", "991K-1.3M", "1.3M-1.65M"))
ggplot(kc_house_data, aes(x=sqft.lot, y=price, fill=sqft.lot)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Land Area") + ylab("Housing Price ($)") + xlab("Land Area (sqft.)") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 
```


Included below is a boxplot of price vs. land area (log version). Log versions are used throughout as a better way to visualize our boxplots (better method than removing outliers).

```{r log price, include=FALSE}
price.ln = log(kc_house_data$price)
```

```{r boxplot lot no outliers, echo=FALSE}
ggplot(kc_house_data, aes(x=sqft.lot, y=price.ln, fill=sqft.lot)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Land Area (log)") + ylab("Housing Price ($)") + xlab("Land Area (sqft.)") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 
```

## Comparison of price and condition

Boxplots of price vs. apartment condition (regular and log versions).

```{r, echo=FALSE}
ggplot(kc_house_data, aes(x=condition, y=price, fill=condition)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Apartment Condition") + ylab("Housing Price ($)") + xlab("Apartment Condition") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 
ggplot(kc_house_data, aes(x=condition, y=price.ln, fill=condition)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Apartment Condition (log)") + ylab("Housing Price ($)") + xlab("Apartment Condition") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 
```

## Comparison of price and grade (construction and design)

Boxplots of price and grade (regular and log versions).

```{r, echo=FALSE}
ggplot(kc_house_data, aes(x=grade, y=price, fill=grade)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Apartment Grade") + ylab("Housing Price ($)") + xlab("Apartment Grade") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 
ggplot(kc_house_data, aes(x=grade, y=price.ln, fill=grade)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Apartment Grade (log)") + ylab("Housing Price ($)") + xlab("Apartment Grade") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 
```


## Comparison of price and # bedrooms

Boxplot of housing price vs. number of bedrooms (log version).

```{r, echo=FALSE}
number.bedrooms <- cut(kc_house_data$bedrooms, breaks = c(0,3,6,9,11), labels = c("1-2", "3-5", "6-8", "9-11"))
ggplot(kc_house_data, aes(x=number.bedrooms, y=price.ln, fill=number.bedrooms)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. # Bedrooms (log)") + ylab("Housing Price ($)") + xlab("# Bedrooms") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 
```

## Comparison of price and year built

Histogram and boxplot of price vs. year built (log version).

```{r, echo=FALSE}
hist(kc_house_data$yr_built, col = heat.colors(12), cex.axis = .75, xlab = "Year Home Built", main = "Histogram of Year Built")
year.built <- cut(kc_house_data$yr_built, breaks = c(1899,1923,1946,1969,1992,2015), labels = c("1900-1923", "1924-1946", "1947-1969", "1970-1992", "1993-2015"))
ggplot(kc_house_data, aes(x=year.built, y=price.ln, fill=year.built)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Year Built (log)") + ylab("Housing Price ($)") + xlab("Year Built") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 
```


# Hypothesis Testing

## Chi-square tests

Below are chi-square tests on condition, grade, bed, and bath. The results for all four are significant (p is close to 0); the variables have a significant relationship with price.

### Chi-square of the condition variable. Significant relationship with price.

H0: The condition and grade are independent.
H1: The condition and grade are not independent.
```{r chisq condition}
cond.tbl <- table(kc_house_data$condition, kc_house_data$grade)
chisq.cond <- chisq.test(cond.tbl)
chisq.cond
```
Since p-value is lower 0.05, we need to reject the null hypothesis. Thus, the condition and grade are not independent. They are corrlated.  

### Chi-square of the "year built" variable. Significant relationship with price.

H0: The year built and grade are independent.
H1: The year built and grade are not independent.

```{r chisq year}
grade.tbl <- table(kc_house_data$grade, year.built)
chisq.cond <- chisq.test(grade.tbl)
chisq.cond
```
Since p-value is lower 0.05, we need to reject the null hypothesis. Thus, the year built and grade are not independent. They are corrlated. 

### Chi-square of the bedroom and bathroom variables. Significant relationship with price.

H0: The price and number of bedrooms are independent.  
H1: The price and number of bedrooms are not independent.  
```{r chisq bedr}
price.intrv = cut(kc_house_data$price, c(0, 250000, 500000, 750000, 1000000, 1250000, 1500000, 1750000, 2000000))
number.bathrooms <- cut(kc_house_data$bathrooms, breaks = c(0,2,4,6,8), labels = c("0.5-1.5", "2-3.5", "4-5.5", "6-8"))
bed_p <- table(number.bedrooms, price.intrv)
x2test1 = chisq.test(bed_p)
x2test1
bath_p <- table(number.bathrooms, price.intrv)
x2test2 = chisq.test(bath_p)
x2test2
```
Since both p-values are lower 0.05, we need to reject the null hypothesis. Thus, the number of bedrooms and the number of bathrooms are not independent from the price of house. They are corrlated. Generally, the price increases when the number of bedrooms and bathrooms become greater.

## Anova tests

Below are ANOVA tests on land area, combined with BP tests and post-hoc Tukey tests.

### ANOVA, BP, and Tukey tests on land area variable.

BP test:  
H0: The variances of prices are the same accross the different sqft of lot levels.  
H1: The variances of prices are different accross the different sqft of lot levels.  

ANOVA test:  
H0: There are no differences between the mean prices of the different sqft.lot levels.  
H1: The mean prices of the different sqft.lot levels are different.  

```{r anova land}
bptest(kc_house_data$price~sqft.lot)
land.anova <- aov(kc_house_data$price~sqft.lot)
summary(land.anova)
TKcond <- TukeyHSD(land.anova)
TKcond
```
Note that the p-value of BP test are 0.996, which is greater than 0.05. Thus, the variances are the same for different groups. Then we use ANOVA test analysis the differences of mean prices. Since the p-value is lower than 0.05, we need to reject the null hypothesis. The mean prices of the different sqft.lot levels are different. From the Turkey test we can see, the p-value of the 660k-991k and 520k-330k, 660k-991k and 330K-660K, 660k-991k and 991K-1.3M are the lowest three. And the p-value for other pairs are very high, which means the means price of 660k-991k(the median level) are higher than others'. Thus the land area of 660k-991k are the most popular level for houses in WA. 


### ANOVA, BP, and Tukey tests on condition variable.

BP test:  
H0: The variances of prices are the same accross different condition.  
H1: The variances of prices are different accross different condition.  

ANOVA test:  
H0: There are no differences between the mean prices of the different condition.  
H1: The mean prices of the different condition are different.  

```{r anova condition}
bptest(price~condition, data = kc_house_data)
cond.anova <- aov(kc_house_data$price~kc_house_data$condition)
summary(cond.anova)
TKcond <- TukeyHSD(cond.anova)
TKcond
```
Note that the p-value of BP test are 0.2923, which is greater than 0.05. Thus, the variances are the same for different condition groups. Then we use ANOVA test analysis the differences of mean prices. Since the p-value is lower than 0.05, we need to reject the null hypothesis. The mean prices of the different condition are different. From the Tukey test we know, the condition 1 and 2 has a very large p-value. Other p-values are very low. Basically, the price increases when the condition is better.   


### ANOVA, BP, and Tukey tests on the living space variable.

H0: The variances of prices are the same accross the different sqft of living levels.  
H1: The variances of prices are different accross the different sqft of living levels.  

```{r anova living}
bptest(kc_house_data$price~sqft.living)
```
Since the p-value is lower than 0.05, we need to reject the null hypothesis. The variances for different sqft of living levels are different. Thus, the ANOVA test is not applicable.  

### ANOVA, BP, and Tukey tests on "year built" variable.

H0: The variances of prices are the same accross the different year of built.  
H1: The variances of prices are different accross the different year of built.  

```{r anova year}
x<-bptest(kc_house_data$price~year.built)
x
```
Since the p-value is lower than 0.05, we need to reject the null hypothesis. The variances for different year built groups are different. Thus, the ANOVA test is not applicable.  


### ANOVA, BP, and Tukey tests on "construction grade" variable.

H0: The variances of prices are the same accross the different grade.  
H1: The variances of prices are different accross the different grade.  

```{r anova grade}
bptest(price~grade, data = kc_house_data)
```
Since the p-value is lower than 0.05, we need to reject the null hypothesis. The variances for different grades are different. Thus, the ANOVA test is not applicable.  

### ANOVA, BP, and Tukey tests on "number of bedrooms" variable.

H0: The variances of prices are the same accross the different number of bedrooms groups.  
H1: The variances of prices are different accross the different number of bedrooms groups.    

```{r anova bedrooms}
bptest(kc_house_data$price~number.bedrooms)
```
Since the p-value is lower than 0.05, we need to reject the null hypothesis. The variances for different number of bedrooms groups are different. Thus, the ANOVA test is not applicable.  

# LSRL


##Model building

Below is our regression model, along with a comprehensive correlation plot.

##First, take a look at all numeric variables and their correlation.
```{r}
h2 <- subset(kc_house_data,select = c(price,bedrooms,bathrooms,sqft_living,sqft_lot,floors,sqft_above,sqft_basement,yr_built ))
hcor = cor(h2)
corrplot(hcor, method = "number", type="upper")
```

##yr_built and sqft_lot seem unrelated to price as their correlation coefficient is almost 0 so I do not chose them as dependent variables to predict the house price.

```{r}
model1 <- lm(price~.-sqft_lot-yr_built,data=h2)
summary(model1)
```

##The coefficient of sqft_basement is NA, which indicates it has a problem with the other variables, so we dropped this one.
And the p-value is bathroom is too large as unsignificant so we dropped this one as well.

```{r}
model2 <- lm(price~.-sqft_lot-yr_built-sqft_basement-bathrooms,data=h2)
summary(model2)
vif(model2)
```

##Everything looks better now; I also checked the vif value of each variable and none of them is too large, it shows that there is no problem of multicollineraty. 
we then added the two factor variables into the dataset to see their effects.

```{r}
h3 <- subset(kc_house_data,select = c(price,bedrooms,sqft_living,sqft_above,floors,grade,condition) )
summary(h3)
```

```{r}
model3 <- lm(price~.,data=h3)
summary(model3)
```

##The 5 levels of condition variable are all unsignificant so we can directly drop condition variable. For the grade variable, the higher grade has significant effect on price prediction. In contrast, low grade does not effect price significantly. 
```{r}
model11 <- lm(price~.-condition,data=h3)
summary(model11)
```
##Now, I added the interaction into the model as I want to see if the correlation of variables would affect the price prediction. I firstly put all interactions into them to see what happen.

```{r}
h4<-subset(kc_house_data,select=c(price,bedrooms,sqft_living,sqft_above,floors,grade))
model4<-lm(price~.+bedrooms:sqft_living+bedrooms:floors+bedrooms:sqft_above+sqft_living:floors+sqft_living:sqft_above+floors:sqft_above,data = h4)
summary(model4)
```
##I drop the unsignificant interactions and some interactions would lead variable to be unsignificant as well so I also drop these ones.Here is what left and this model seems nice. 

```{r digit=9}
model5<-lm(price~.+bedrooms:sqft_above+sqft_living:sqft_above,data = h4)
summary(model5)
```

##Price = 142000 +bedrooms*(-31710+10.72*sqft_above)+sqft_living*(170+2.943*sqft_above)+sqft_above*(-228.6)+floors*14570+grade()*
Here is our final model to predict the house price. 
##Problem: As the price histogram above is quite left-skewed, it means there are many outliers whose price is very high in the dataset.While I do the model building, I do not exclude the outliers as I think these values are also important. So our final model is also skewed little. It means that for low price hoouse, our model may predict higher price than true value, but for high price house, our model will predict lower than actual one. 

