---
title: "6101 pres"
author: "Changhao Ying"
date: "2019Äê11ÔÂ16ÈÕ"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##PCA#
```{r include=FALSE}
library("dplyr")
library("tidyr")
library("tidyverse")
library("ggplot2")
library("Rmisc")
library("MASS")
library("corrplot")
library("faraway")
library("factoextra")
library("pls")
library("magrittr")
```

```{r}
house <- data.frame(read.csv("kc_house_data.csv", header = TRUE))
str(house)
```

##subset data 
```{r}
house2 <- subset(house,select = -c(id,date,waterfront,view,zipcode,lat,long,sqft_living15,sqft_lot15,yr_renovated))
str(house2)
```
##All numeric variables. 

```{r}
sum(is.na(house2))
```



```{r}
corx=cor(house2[,-1])
corx
```



##Here we have 10 variables which are correlated and we want to reduce the number of them, pick up the important ones.

##PCA part,scale=T normalize the variables becuase they have different scales.
```{r}
prin_comp <- prcomp(scale(house2[,-1]))
summary(prin_comp)
prin_comp$center ##output the mean of variables
prin_comp$scale ##output the sd of variables
```

```{r}
prin_comp$rotation
```

```{r}
fviz_eig(prin_comp)
```





```{r}
biplot(prin_comp, scale = 0)
##The first principle component looks like sqft_above, the second looks like sqft_basement)
```

```{r}
biplot(prin_comp,6:7, scale =0)
```


```{r}
pr.var <- (prin_comp$sdev^2)
pve <- pr.var/sum(pr.var)
plot(cumsum(pve), xlab="Principal Component (standardized)", ylab ="Cumulative Proportion of Variance Explained",ylim=c(0,1),type="b")
```

##7 components result in close to 95% 


##Predicting use pca

```{r}
houseprice <- prin_comp$x
modHouses <- lm(house2$price ~ houseprice[,1:7])
summary(modHouses)
```

```{r}
fullmodel=lm(price~.-sqft_basement,data=house2)
summary(fullmodel)
```


```{r}
par(mfrow = c(1,2))
plot(house2$price, predict(modHouses), xlab = "actual price", ylab = "Predicted price", main = "PCR", abline(a = 0, b = 1, col = "red"))
plot(house2$price, predict(fullmodel), xlab = "actual price", ylab = "Predicted price", main = "Full model", abline(a = 0, b = 1, col = "red"))
```

##R^2 of full model is 0.618 which is higher than pcr which is 0.613
##we can see that both models underestimate the value of house price over 4e+06



