---
title: "Housing Market - Seattle Area"
author: "Anti-Code Group"
date: "10/5/2019"
output:  
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(scipen = 999)
```

```{r basicfunct, include=FALSE}
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
```

```{r base_lib, include=FALSE}
loadPkg("dplyr")
loadPkg("tidyr")
loadPkg("tidyverse")
loadPkg("ggplot2")
loadPkg('ggmap')
loadPkg('rjson')
loadPkg('jsonlite')
loadPkg('leaflet')
loadPkg('grDevices')
loadPkg('scales')
loadPkg('RCurl')
loadPkg('sp')
loadPkg('geojsonio')
```

# Description Data
## Source Data

```{r, echo=FALSE}
kc_house_data <- read.csv("kc_house_data.csv")
str(kc_house_data)
colnames(kc_house_data)
```

```{r clean, include=FALSE}
kc_house_data <- subset(kc_house_data, select = -c(9, 10))

kc_house_data <- subset(kc_house_data, kc_house_data$bedrooms != 0)

kc_house_data <- subset(kc_house_data, kc_house_data$bathrooms != 0)

kc_house_data <- subset(kc_house_data, kc_house_data$bedrooms < 30)

kc_house_data <-  drop_na(kc_house_data)

kc_house_data$condition <- as.factor(kc_house_data$condition)

kc_house_data$grade <- as.factor(kc_house_data$grade)

```

## Summary for Price
```{r, echo=FALSE}
summary(kc_house_data$price)

print(mean(kc_house_data$price))

print(sd(kc_house_data$price))

print(var(kc_house_data$price))

```

## Geographic Coverage of Data
Below is a visualization of the points in the dataset by price, plotted with the leaflet library.
 Note that the data have been divided by quantiles to provide a better understanding of the distribution of rental price.
 
```{r geo, echo=FALSE}
price.bins <- c(0, 250000, 500000, 750000, 1000000, 1250000, 1500000, 1750000, 2000000, 8000000)
qpal <-  colorBin(palette = 'GnBu', kc_house_data$price, bins= price.bins, n = 9)
house.map <- leaflet(kc_house_data) %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addCircleMarkers(lng = ~long, lat = ~lat, 
                   stroke = FALSE, 
                   fillOpacity = 5, 
                   color = ~qpal(price), 
                   radius = 2,
                   label = ~as.character(paste0('Price: $', price, ", ", 'condition: ', condition, ", ", 'year built: ', yr_built))) %>%
  addLegend('bottomright', pal = qpal, values = ~price, opacity = 1, title = 'Price', labFormat = labelFormat(prefix = '$', between = '- $'))

house.map

```

## Distribution of Price

The following are housing price distributions (histograms). The first includes the frequency of all price occurences, and the other only shows those less than or equal to $2M (the vast majority of data points are below this price point). Also included is a Q-Q plot showing the distribution of the data.

```{r price hist, echo=FALSE}

hist(kc_house_data$price, xaxt="n", ylim = c(0,14000), col = "green", main = "Housing Price Histogram", xlab = "Housing Price ($)", cex.axis = .75)
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))


hist(kc_house_data$price[kc_house_data$price<=2000000], xaxt="n", ylim = c(0,5000), col = heat.colors(20), main = "Housing Price Histogram, $0-$2M Only", xlab = "Housing Price ($)", cex.axis = .75)
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))

qqnorm(kc_house_data$price, pch = 20, cex = .5, main = "Housing Prices: Normal Q-Q Plot", ylab = "Housing Price ($)", cex.axis = .75)

```

##The following is a summary of price per square foot, per annum.

```{r price per year}

date_price <- kc_house_data[c('date', 'price', 'sqft_living', 'sqft_lot')]
date_price$date <- as.character(date_price$date)
date_price$date <- substr(date_price$date, 1, nchar(date_price$date)-11)
date_price <- date_price[order(date_price$date),]
date_price$date <- as.factor(date_price$date)
date_price %>%
  group_by(date) %>%
  summarise(price = sum(price), sqft_living=sum(sqft_living), sqft_lot=sum(sqft_lot), price_per_sqft = price/sqft_living)

```

price per sqft:  
  2014 -> 36.06  
  2015 -> 35.11  

  
## High observations

The following are a summary of the first few rows for bedrooms, land area, and living space (price showed is average for that variable level (e.g. $262,000 is the average price for grade 3 construction).

```{r top price, echo=FALSE}
print(top_n(kc_house_data, 5, price)) %>% arrange(desc(price))
```

```{r top sqftliv, echo=FALSE}
print(top_n(kc_house_data, 5, sqft_living)) %>% arrange(desc(sqft_living))
```

```{r top sqftlot, echo=FALSE}
print(top_n(kc_house_data, 5, sqft_lot))  %>% arrange(desc(sqft_lot))
```


```{r top bedrooms, echo=FALSE}
print(top_n(kc_house_data, 4, bedrooms))  %>% arrange(desc(bedrooms))
```

```{r rop avg cond, echo=F}
print(aggregate(price~condition,data=kc_house_data,FUN=mean))

```

```{r rop avg grade, echo=F}
print(aggregate(price~grade,data=kc_house_data,FUN=mean))

```


## Comparison of price and living sqft

Below is a scatterplot of price vs. living space, along with their correlation.

```{r plot sqft living, echo=FALSE}

plot(kc_house_data$sqft_living, kc_house_data$price, xaxt="n", yaxt="n", pch = 20, cex = .1, xlab = "Square Feet of Living Space", ylab = "Housing Price ($)", cex.axis = .75, main = "Housing Price vs. Living Space")
abline(lm(kc_house_data$price ~ kc_house_data$sqft_living, data = kc_house_data), col = "orange")
legend(x='topright', legend=paste('Correlation =',round(cor(kc_house_data$sqft_living, kc_house_data$price),5)))
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))
axis(side=2, at=axTicks(2), 
     labels=formatC(axTicks(2), format="d", big.mark=','))
```

## Boxplot & Scatterplot for Living Space

Below are a boxplot and scatterplot for living space vs. price.

```{r boxplot sqft living, echo=FALSE}

max(kc_house_data$sqft_living)
min(kc_house_data$sqft_living)

sqft.living <- cut(kc_house_data$sqft_living, breaks = c(289,2940,5590,8240,10890,13540), labels = c("290-2,940", "2,941-5,590", "5,591-8,240", "8,241-10,890", "10,891-13,540"))


ggplot(kc_house_data, aes(x=sqft.living, y=price, fill=sqft.living)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Living Space") + ylab("Housing Price ($)") + xlab("Square Feet of Living Space") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

```


## Comparison price and lot sqft

The following includes histograms of lot square feet (including one equal to or below 25,000 sqft., where the vast majority of data points lie). Also included is a scatterplot showing the correlation between price and land area.

```{r lot summary, echo=FALSE}
summary(kc_house_data$sqft_lot)

```

```{r plot lot, echo=FALSE}

hist(kc_house_data$sqft_lot, col = "yellow", cex.axis = .75, xlab = "Land Area (sqft.)", main = "Histogram of Land Area")

hist(kc_house_data$sqft_lot[kc_house_data$sqft_lot<=25000], xaxt="n", cex.axis = .75, xlab = "Land Area (sqft.)", main = "Histogram of Land Area, up to 25,000 sqft.", col = heat.colors(13))
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))

plot(kc_house_data$sqft_lot,kc_house_data$price, cex.axis = .75, pch = 20, cex = .1, xlab = "Land Area (sqft.)", ylab = "Housing Price ($)", xaxt="n", yaxt="n", main = "Housing Price vs. Land Area")
abline(lm(kc_house_data$price~kc_house_data$sqft_lot, data = kc_house_data), col = "orange")
legend(x='topright', legend=paste('Correlation =',round(cor(kc_house_data$sqft_lot, kc_house_data$price),5)))
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))
axis(side=2, at=axTicks(2), 
     labels=formatC(axTicks(2), format="d", big.mark=','))

```

## Comparison of Land Area and Housing Price

Included below is a boxplot of land area vs. housing price.

```{r boxplot lot, echo=FALSE}

max(kc_house_data$sqft_lot)
min(kc_house_data$sqft_lot)

sqft.lot <- cut(kc_house_data$sqft_lot, breaks = c(519,330688,660856,991023,1321191, 1651359), labels = c("520-330K", "330K-660K", "660K-991K", "991K-1.3M", "1.3M-1.65M"))

ggplot(kc_house_data, aes(x=sqft.lot, y=price, fill=sqft.lot)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Land Area") + ylab("Housing Price ($)") + xlab("Land Area (sqft.)") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 


```
## Comparison of Price vs. Land Area

Included below is a boxplot of price vs. land area (log version).

```{r log price, include=FALSE}
price.ln = log(kc_house_data$price)

```

```{r boxplot lot no outliers, echo=FALSE}

ggplot(kc_house_data, aes(x=sqft.lot, y=price.ln, fill=sqft.lot)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Land Area (log)") + ylab("Housing Price ($)") + xlab("Land Area (sqft.)") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

```

## Comparison of price and condition

Boxplots of price vs. apartment condition (regular and log versions).

```{r, echo=FALSE}

ggplot(kc_house_data, aes(x=condition, y=price, fill=condition)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Apartment Condition") + ylab("Housing Price ($)") + xlab("Apartment Condition") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 


ggplot(kc_house_data, aes(x=condition, y=price.ln, fill=condition)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Apartment Condition (log)") + ylab("Housing Price ($)") + xlab("Apartment Condition") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

```

## Comparison of price and grade (construction and design)

Boxplots of price and grade (regular and log versions).

```{r, echo=FALSE}

ggplot(kc_house_data, aes(x=grade, y=price, fill=grade)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Apartment Grade") + ylab("Housing Price ($)") + xlab("Apartment Grade") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

ggplot(kc_house_data, aes(x=grade, y=price.ln, fill=grade)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Apartment Grade (log)") + ylab("Housing Price ($)") + xlab("Apartment Grade") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

```


## Comparison of price and # bedrooms

Boxplot of housing price vs. number of bedrooms (log version).

```{r, echo=FALSE}

number.bedrooms <- cut(kc_house_data$bedrooms, breaks = c(0,3,6,9,11), labels = c("1-2", "3-5", "6-8", "9-11"))

ggplot(kc_house_data, aes(x=number.bedrooms, y=price.ln, fill=number.bedrooms)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. # Bedrooms (log)") + ylab("Housing Price ($)") + xlab("# Bedrooms") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

```

## Comparison of price and year built

Histogram and boxplot of price vs. year built (log version).

```{r, echo=FALSE}

hist(kc_house_data$yr_built, col = heat.colors(12), cex.axis = .75, xlab = "Year Home Built", main = "Histogram of Year Built")

year.built <- cut(kc_house_data$yr_built, breaks = c(1899,1923,1946,1969,1992,2015), labels = c("1900-1923", "1924-1946", "1947-1969", "1970-1992", "1993-2015"))

ggplot(kc_house_data, aes(x=year.built, y=price.ln, fill=year.built)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Year Built (log)") + ylab("Housing Price ($)") + xlab("Year Built") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

```

# Hypothesis Testing
## Anova tests

ANOVA run on price and living space; Tukey post-hoc test also conducted.

```{r anova sqfliv}

living.anova <- aov(kc_house_data$price~sqft.living)
summary(living.anova)

TKcond <- TukeyHSD(living.anova)
TKcond
```

ANOVA run on price and land area; Tukey post-hoc test also conducted.

```{r anova land}

land.anova <- aov(kc_house_data$price~sqft.lot)
summary(land.anova)

TKcond <- TukeyHSD(land.anova)
TKcond
```

ANOVA run on price and year built; Tukey post-hoc test also conducted.

```{r anova year}

year.anova <- aov(kc_house_data$price~year.built)
summary(year.anova)

TKcond <- TukeyHSD(year.anova)
TKcond
```

ANOVA run on price and number of bedrooms; Tukey post-hoc test also conducted.

```{r anova bedrooms}

beds.anova <- aov(kc_house_data$price~number.bedrooms)
summary(beds.anova)

TKcond <- TukeyHSD(beds.anova)
TKcond
```


## Chi-square tests

Chi-square test run on price and housing condition.

```{r chisq condition}

cond.tbl <- table(kc_house_data$price, kc_house_data$condition)
chisq.cond <- chisq.test(cond.tbl)
chisq.cond
```

Chi-square test run on price and grade.

```{r anova grade}

grade.tbl <- table(kc_house_data$price, kc_house_data$grade)
chisq.grade <- chisq.test(grade.tbl)
chisq.grade
```



