---
title: "Housing Market - Seattle Area"
author: "Anti-Code Group"
date: "10/5/2019"
output:  
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(scipen = 999)
```

```{r basicfunct, include=FALSE}
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
```

```{r base_lib, include=FALSE}
loadPkg("dplyr")
loadPkg("tidyr")
loadPkg("tidyverse")
loadPkg("ggplot2")
loadPkg('ggmap')
loadPkg('rjson')
loadPkg('jsonlite')
loadPkg('leaflet')
loadPkg('grDevices')
loadPkg('scales')
loadPkg('RCurl')
loadPkg('sp')
loadPkg('geojsonio')
loadPkg('lmtest')
loadPkg("faraway")
loadPkg("corrplot")
loadPkg("modelr")
```

# Chapter 1: Introduction

*insert intro here: why this dataset, background research on topic and data, what this data and report about, table of contents description*

Seattle boasts among the "hottest" housing markets in the United States; as of July 2018, Seattle "led the nation in home price gains" for 21 straight months, topped only by Portland in the 1990s – a trend driven by the city's tech sector and a  lack of supply compared with demand (https://bit.ly/2v5UMcn). Given the Seattle housing market's notoriety for high prices, we were interested in exploring which variables affect housing price in this market. With this goal in mind, we found a public dataset on Kaggle ("House Sales in King County, USA") offering 21,613 observations across 21 variables. Although it doesn't explore macro-level variables affecting housing price (such as the local job market, Amazon presence, etc.), it does focus on micro-level variables, such as renovations, number of bedrooms, square feet of living space, etc. that are common to virtually all housing markets in the United States. An analysis of Seattle's housing market could lay the groundwork for future comparative analysis with other housing markets.

This report is organized as follows:

1. Description of the Data (explanation of the dataset and its variables, 
2. Geographic Coverage of Data
3. House Prices vs. Size
4. Factors Analysis
5. Multiple Linear Regression Model
6. Conclusion

# Chapter 2: Description Data
## Source Data

https://www.kaggle.com/harlfoxem/housesalesprediction/data

https://github.com/matteobucalossi50/DATS6101_Project

*can delete links and explain dataset and variables and all that, as well as what done with cleaning coding (including reasoning for log price, whici will later be more clear with graphs)* 

As mentioned previously, our dataset houses 21,613 observations across 21 variables. (See below for a readout of the dataset's structure and variable names.) Variable descriptions are as follows and come from the following link: https://bit.ly/2MsyRFl; astericks next to variable name indicates usage in our analysis:

```{r, echo=FALSE}
kc_house_data <- read.csv("kc_house_data.csv")
str(kc_house_data)
colnames(kc_house_data)
```

1. Id (unique ID for each home sold)
2. Date (date of the home sale)
3. Price (price of each home sold)***
4. Bedrooms (number of bedrooms)***
5. Bathrooms (number of bathrooms, where .5 accounts for a room with a toilet but no shower)
6. Sqft_living (square footage of the apartments' interior living space)***
7. Sqft_lot (square footage of the land space)***
8. Floors (number of floors)
9. Waterfront (a dummy variable for whether the apartment was overlooking the waterfront or not; 0 represents no waterfront)
10. View (an index from 0 to 4 of how the view of the property was)
11. Condition (an index from 1 to 5 on the condition of the apartment, with the lowest number representing poor condition)***
12. Grade (an index from 1 to 13, with the lowest number representation poor construction and design)***
13. Sqft_above (the the square footage of the interior housing space that is above ground level)
14. Sqft_basement (the square footage of the interior housing space that is below ground level)
15. yr_built (the year the house was initially built)***
16. yr_renovated (the year of the house's last renovation)***
17. zipcode (what zipcode area the house is in)***
18. Lat (latitude)***
19. Long (longitude)***
20. Sqft_living15 (the square footage of the interior housing living space for the nearest 15 neighbors)
21. Sqft_lot15 (the square footage of the land lots of the nearest 15 neighbors)

We ignored "Id" and "Date" because these are independent variables with no relation to price. We also ignored "floors" because it can be considered a proxy for sqft_living. "Waterfront" and "View" were dropped because the vast majority of properties were coded as "0". We ignored sqft_basement and sqft_above because they were corollaries of "sqft_living" (we didn't want redundancy in our analysis). We also ignored "sqft_living 15" and "sqft_lot15" because we were interested only in the attributes of individual houses, not those of their surrounding neighborhoods (although that could make for an interesting follow-up study).

Following these decisions, we cleaned the data accordingly: we dropped "waterfront" and "view"; we subsetted the dataset to include only properties with more than 0 bedrooms and bathrooms (we considered these "outlier" properties); we subsetted the dataset to include only properties with less than 30 bedrooms (again, to exclude outlier properties); we dropped "NA" values from the dataset to simplify our analysis ("NA" values are hard to perform operations on); we converted "condition" and "grade" into factor variables because they are effectively intervals; and we ran "housing price" through a logarithmic function to make for better visualization.

```{r clean, include=FALSE}
kc_house_data <- subset(kc_house_data, select = -c(9, 10))

kc_house_data <- subset(kc_house_data, kc_house_data$bedrooms != 0)

kc_house_data <- subset(kc_house_data, kc_house_data$bathrooms != 0)

kc_house_data <- subset(kc_house_data, kc_house_data$bedrooms < 30)

kc_house_data <-  drop_na(kc_house_data)

kc_house_data$condition <- as.factor(kc_house_data$condition)

kc_house_data$grade <- as.factor(kc_house_data$grade)

price.ln = log(kc_house_data$price)

```

```{r, include=FALSE}
summary(kc_house_data$price)

print(mean(kc_house_data$price))

print(sd(kc_house_data$price))

print(var(kc_house_data$price))

```

*use inline code to describe dataset statistics and values*

## Geographic Coverage of Data
Below is a visualization of the points in the dataset by price, plotted with the leaflet library.
Note that the data have been divided by unequal bins to provide a better visualization of the distribution of housing price, so please read the legend carefully.
 
```{r geo, echo=FALSE}
price.bins <- c(0, 250000, 500000, 750000, 1000000, 1250000, 1500000, 1750000, 2000000, 8000000)
qpal <-  colorBin(palette = 'GnBu', kc_house_data$price, bins= price.bins, n = 9)
house.map <- leaflet(kc_house_data) %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addCircleMarkers(lng = ~long, lat = ~lat, 
                   stroke = FALSE, 
                   fillOpacity = 5, 
                   color = ~qpal(price), 
                   radius = 2,
                   label = ~as.character(paste0('Price: $', price, ", ", 'condition: ', condition, ", ", 'year built: ', yr_built))) %>%
  addLegend('bottomright', pal = qpal, values = ~price, opacity = 1, title = 'Price', labFormat = labelFormat(prefix = '$', between = ' - $'))

house.map

```

Here instead is a visualization of the observations by property lot sqft. Again, data have been divided by unequal bins to provide a better visualization of the distribution of housing price, so please read the legend carefully.

```{r geo, echo=FALSE}

sqft.bins <- c(519,10000,200000,400000,660856, 1651359)

qpal <-  colorBin(palette = 'YlOrBr', kc_house_data$sqft_lot, bins= sqft.bins, n = 5)
house.map <- leaflet(kc_house_data) %>% 
  addProviderTiles("CartoDB.DarkMatter") %>% 
  addCircleMarkers(lng = ~long, lat = ~lat, 
                   stroke = FALSE, 
                   fillOpacity = 5, 
                   color = ~qpal(sqft_lot), 
                   radius = 2,
                   label = ~as.character(paste0('Sqft_lot: $', sqft_lot, ", ", 'condition: ', condition, ", ", 'year built: ', yr_built, 'price: ', price))) %>%
  addLegend('bottomright', pal = qpal, values = ~sqft_lot, opacity = 1, title = 'Sqft_lot', labFormat = labelFormat( between = ' - '))

house.map

```



# Chapter 3: House prices and sizes
## House Price

*descibe how price is min max etc, then inline code for price per sqft and how expensive seattle is etc*

```{r price hist, echo=FALSE}

hist(kc_house_data$price, xaxt="n", ylim = c(0,14000), col = "green", main = "Housing Price Histogram", xlab = "Housing Price ($)", cex.axis = .75)
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))


hist(kc_house_data$price[kc_house_data$price<=2000000], xaxt="n", ylim = c(0,5000), col = heat.colors(20), main = "Housing Price Histogram, $0-$2M Only", xlab = "Housing Price ($)", cex.axis = .75)
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))

qqnorm(kc_house_data$price, pch = 20, cex = .5, main = "Housing Prices: Normal Q-Q Plot", ylab = "Housing Price ($)", cex.axis = .75)

```

```{r price per year, include=FALSE}

date_price <- kc_house_data[c('date', 'price', 'sqft_living', 'sqft_lot')]
date_price$date <- as.character(date_price$date)
date_price$date <- substr(date_price$date, 1, nchar(date_price$date)-11)
date_price <- date_price[order(date_price$date),]
date_price$date <- as.factor(date_price$date)
date_price %>%
  group_by(date) %>%
  summarise(price = sum(price), sqft_living=sum(sqft_living), sqft_lot=sum(sqft_lot), price_per_sqft = price/sqft_living)

```

price per sqft:  
  2014 -> 36.06  
  2015 -> 35.11  

### how do the most expensive houses look like?

```{r top price, echo=FALSE}
print(top_n(kc_house_data, 5, price)) %>% arrange(desc(price))
```

### what is the average price for each condition level?
```{r rop avg cond, echo=F}
print(aggregate(price~condition,data=kc_house_data,FUN=mean))

```

### what is the average price for each grade level?
```{r rop avg grade, echo=F}
print(aggregate(price~grade,data=kc_house_data,FUN=mean))

```

## House Size


*add histograms for bedroom variable and sqft living variable - comment on them, why such house be that in real world etc etc..*

```{r lot hist, echo=FALSE}
hist(kc_house_data$sqft_lot, col = "yellow", cex.axis = .75, xlab = "Land Area (sqft.)", main = "Histogram of Land Area")

hist(kc_house_data$sqft_lot[kc_house_data$sqft_lot<=25000], xaxt="n", cex.axis = .75, xlab = "Land Area (sqft.)", main = "Histogram of Land Area, up to 25,000 sqft.", col = heat.colors(13))
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))
```

### How do the largest houses look like?
*introduce the reasoning of this passage and thendescribe each output with reason under each of them*
```{r top sqftliv, echo=FALSE}
print(top_n(kc_house_data, 5, sqft_living)) %>% arrange(desc(sqft_living))
```


```{r top sqftlot, echo=FALSE}
print(top_n(kc_house_data, 5, sqft_lot))  %>% arrange(desc(sqft_lot))
```


```{r top bedrooms, echo=FALSE}
print(top_n(kc_house_data, 4, bedrooms))  %>% arrange(desc(bedrooms))
```


# Chapter 4: Factors Analysis
## SMART Question
Are houses of different sizes priced differently?

### Comparison of sqft living with price

```{r plot sqft living, echo=FALSE}

plot(kc_house_data$sqft_living, kc_house_data$price, xaxt="n", yaxt="n", pch = 20, cex = .1, xlab = "Square Feet of Living Space", ylab = "Housing Price ($)", cex.axis = .75, main = "Housing Price vs. Living Space")
abline(lm(kc_house_data$price ~ kc_house_data$sqft_living, data = kc_house_data), col = "orange")
legend(x='topright', legend=paste('Correlation =',round(cor(kc_house_data$sqft_living, kc_house_data$price),5)))
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))
axis(side=2, at=axTicks(2), 
     labels=formatC(axTicks(2), format="d", big.mark=','))
```

```{r boxplot sqft living, echo=FALSE}

max(kc_house_data$sqft_living)
min(kc_house_data$sqft_living)

sqft.living <- cut(kc_house_data$sqft_living, breaks = c(289,2940,5590,8240,10890,13540), labels = c("290-2,940", "2,941-5,590", "5,591-8,240", "8,241-10,890", "10,891-13,540"))


ggplot(kc_house_data, aes(x=sqft.living, y=price, fill=sqft.living)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Living Space") + ylab("Housing Price ($)") + xlab("Square Feet of Living Space") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

```

*explains bp test with inline code and not possible anova*
```{r anova living, include=FALSE}
bptest(kc_house_data$price~sqft.living)

```

### Comparison of sqft lot with price

```{r lot summary, include=FALSE}
summary(kc_house_data$sqft_lot)

```

```{r plot lot, echo=FALSE}

plot(kc_house_data$sqft_lot,kc_house_data$price, cex.axis = .75, pch = 20, cex = .1, xlab = "Land Area (sqft.)", ylab = "Housing Price ($)", xaxt="n", yaxt="n", main = "Housing Price vs. Land Area")
abline(lm(kc_house_data$price~kc_house_data$sqft_lot, data = kc_house_data), col = "orange")
legend(x='topright', legend=paste('Correlation =',round(cor(kc_house_data$sqft_lot, kc_house_data$price),5)))
axis(side=1, at=axTicks(1), 
     labels=formatC(axTicks(1), format="d", big.mark=','))
axis(side=2, at=axTicks(2), 
     labels=formatC(axTicks(2), format="d", big.mark=','))

```

```{r boxplot lot, echo=FALSE}

max(kc_house_data$sqft_lot)
min(kc_house_data$sqft_lot)

sqft.lot <- cut(kc_house_data$sqft_lot, breaks = c(519,330688,660856,991023,1321191, 1651359), labels = c("520-330K", "330K-660K", "660K-991K", "991K-1.3M", "1.3M-1.65M"))

ggplot(kc_house_data, aes(x=sqft.lot, y=price, fill=sqft.lot)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Land Area") + ylab("Housing Price ($)") + xlab("Land Area (sqft.)") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 


```

```{r boxplot lot log, echo=FALSE}

ggplot(kc_house_data, aes(x=sqft.lot, y=price.ln, fill=sqft.lot)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Land Area (log)") + ylab("Housing Price ($)") + xlab("Land Area (sqft.)") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

```

*explain results of anova inline code etc*

```{r anova land}

bptest(kc_house_data$price~sqft.lot)

land.anova <- aov(kc_house_data$price~sqft.lot)
summary(land.anova)

TKcond <- TukeyHSD(land.anova)
TKcond
```


### Comparison of number of bedrooms with price
```{r, echo=FALSE}

number.bedrooms <- cut(kc_house_data$bedrooms, breaks = c(0,3,6,9,11), labels = c("1-2", "3-5", "6-8", "9-11"))

ggplot(kc_house_data, aes(x=number.bedrooms, y=price.ln, fill=number.bedrooms)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. # Bedrooms (log)") + ylab("Housing Price ($)") + xlab("# Bedrooms") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

```

*same for anova bedrooms*

```{r anova bedrooms}
bptest(kc_house_data$price~number.bedrooms)

```

*can use chisq to see if more rooms more cost - smart question to intro, explain chisq (hypothesis, categ tranformation, results etc*
```{r chisq rooms}

price.intrv = cut(kc_house_data$price, c(0, 250000, 500000, 750000, 1000000, 1250000, 1500000, 1750000, 2000000))

number.bathrooms <- cut(kc_house_data$bathrooms, breaks = c(0,2,4,6,8), labels = c("0.5-1.5", "2-3.5", "4-5.5", "6-8"))

bed_p <- table(number.bedrooms, price.intrv)
x2test1 = chisq.test(bed_p)
x2test1

bath_p <- table(number.bathrooms, price.intrv)
x2test2 = chisq.test(bath_p)
x2test2

```



## SMART Question
Are houses of different quality(?) priced differently? *check all these questions make sense given what nick said*

### Comparison of condition with price
**explain well this variable, what represents etc as they asked*

```{r, echo=FALSE}

ggplot(kc_house_data, aes(x=condition, y=price, fill=condition)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Apartment Condition") + ylab("Housing Price ($)") + xlab("Apartment Condition") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 


ggplot(kc_house_data, aes(x=condition, y=price.ln, fill=condition)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Apartment Condition (log)") + ylab("Housing Price ($)") + xlab("Apartment Condition") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

```

*explains anova test tukey and all*
```{r anova condition}
bptest(price~condition, data = kc_house_data)

cond.anova <- aov(kc_house_data$price~kc_house_data$condition)
summary(cond.anova)

TKcond <- TukeyHSD(cond.anova)
TKcond
```

### Comparison of grade with price 
*explain well this variable, what represents etc as they asked, design/construction*

```{r, echo=FALSE}

ggplot(kc_house_data, aes(x=grade, y=price, fill=grade)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Apartment Grade") + ylab("Housing Price ($)") + xlab("Apartment Grade") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

ggplot(kc_house_data, aes(x=grade, y=price.ln, fill=grade)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Apartment Grade (log)") + ylab("Housing Price ($)") + xlab("Apartment Grade") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

```

*no anova..*
```{r anova grade, include=FALSE}
bptest(price~grade, data = kc_house_data)

```


*explains chi-sq results and how they related*
```{r chisq grade}

cond.tbl <- table(kc_house_data$condition, kc_house_data$grade)
chisq.cond <- chisq.test(cond.tbl)
chisq.cond
```


## SMART Question
Are older houses priced differently?
### Comparison of year they were built with price

```{r, echo=FALSE}

hist(kc_house_data$yr_built, col = heat.colors(12), cex.axis = .75, xlab = "Year Home Built", main = "Histogram of Year Built")

year.built <- cut(kc_house_data$yr_built, breaks = c(1899,1923,1946,1969,1992,2015), labels = c("1900-1923", "1924-1946", "1947-1969", "1970-1992", "1993-2015"))

ggplot(kc_house_data, aes(x=year.built, y=price.ln, fill=year.built)) + geom_boxplot() + scale_fill_brewer(palette="Spectral") + ggtitle("Housing Price vs. Year Built (log)") + ylab("Housing Price ($)") + xlab("Year Built") +  theme(plot.title= element_text(hjust=0.5, size = 14)) + scale_y_continuous(labels = comma) 

```

*no anova..*
```{r anova year}
bptest(kc_house_data$price~year.built)

```

*explain results and problems with it, introduce for next analysis below*

### Comparison of year renovated with price

*charts and all*

*do anova for this one too*


**compared renovated with built and see price changes too..**


### Are older houses worst in terms of quality?
*use chisq to assess old and quality independence*

```{r chisq year}

gradey.tbl <- table(kc_house_data$grade, year.built)
chisq.gradey <- chisq.test(gradey.tbl)
chisq.gradey

```

```{r chisq year}

condy.tbl <- table(kc_house_data$condition, year.built)
chisq.condy <- chisq.test(condy.tbl)
chisq.condy

```

### Does quality get better when renovated?
*use chisq to assess renovated and quality indep - replicate above chisq tests for year.renov*
 


**subset for older houses with renovation and rerun both tests and see difference in time**

like older house renovated cost normal? median house are still not renovated? how is the condition grade for older one? how is it for older renovated how is it for newer? how price changes in all these dynamics etc
like they wanted us to go in depth of that cause it actually makes sense


# Chapter 5: Multiple Linear Regression Model

## SMART Question:
What factors influence the house price the most?

### LSRL Model building

Below is our regression model, along with a comprehensive correlation plot.
```{r}
h2 <- subset(kc_house_data,select = c(price,bedrooms,bathrooms,sqft_living,sqft_lot,floors,sqft_above,sqft_basement,yr_built ))
hcor = cor(h2)
corrplot(hcor, method = "number", type="upper")
```

yr_built and sqft_lot seem unrelated to price.

```{r}
model1 <- lm(price~.-sqft_lot-yr_built,data=h2)
summary(model1)
```

The coefficient of sqft_basement is NA indicates it has problem with other variables so I drop this one.

```{r}
model2 <- lm(price~.-sqft_lot-yr_built-sqft_basement-bathrooms,data=h2)
summary(model2)
vif(model2)
```

Everything looks better, then I added the two factor variables into the dataset and se their effects.

```{r}
h3 <- subset(kc_house_data,select = c(price,bedrooms,sqft_living,sqft_above,floors,grade,condition) )
summary(h3)
```

```{r}
model3 <- lm(price~.,data=h3)
summary(model3)
vif(model3)
```
```{r}
model11 <- lm(price~.-condition,data=h3)
summary(model11)
vif(model11)
```

All levels of condition are unsignificant, so we drop the condition varibale. And high grade has significant effect on price.

```{r}
h4<-subset(kc_house_data,select=c(price,bedrooms,sqft_living,sqft_above,floors,grade))
model4<-lm(price~.+bedrooms:sqft_living+bedrooms:floors+bedrooms:sqft_above+sqft_living:floors+sqft_living:sqft_above+floors:sqft_above,data = h4)
summary(model4)
vif(model4)
```
```{r digit=9}
model5<-lm(price~.+bedrooms:sqft_above+sqft_living:sqft_above,data = h4)
summary(model5)
vif(model5)
```

### Final results and approved model for preidiction of price
*explain coefficients and all that - use inline coding etc..*


# Chapter 6: Conclusion
*final insights, main relationships, predictors, what to look in a real estate dataset etc..*
*future openings for further studies, analysis, tests etc on this..*


# Bibliography

